---
title: "Public Health Whole Genome Sequencing Analysis Report"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    navbar:
      - { title: "ODHL", href:
      "https://odh.ohio.gov/know-our-programs/microbiology/microbiology",
      align_right}
    logo: "odh_logo_240128.png"
params:
  configFILE: "REP_CONFIG"
  sampletable: "REP_SAMPLETABLE"
  arpredictions: "REP_PREDICT"
  projectname: "REP_PROJID"
  date: "REP_DATE"
  testing: "N"
editor_options: 
  chunk_output_type: console
---

```{r libs, echo=FALSE,include=FALSE,warning=FALSE,message=FALSE}
library(ape)
library(argparser)
library(crosstalk)
library(DT)
library(flexdashboard)
library(ggplot2)
library(ggpubr)
library(heatmaply)
library(htmlwidgets)
library(kableExtra)
library(pander)
library(phytools)
library(plotly)
library(RColorBrewer)
library(rgl)
library(rmarkdown)
library(yaml)
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE)
options(rgl.useNULL = TRUE) # Suppress the separate window.
'%ni%' <- function(x,y)!('%in%'(x,y))
```

```{r args, echo=FALSE,warning=FALSE}
username="S. Chill"
reporttype="Standard"

#position args
if (params$testing=="N"){
  projectname=params$projectname
  configFILE=params$configFILE
  date=params$date
  sampletable=params$sampletable
  arpredictions=params$arpredictions
} else {
  projectname="OH-M6588-231201"
  outdir="OH-M6588-231201/"
  date="12/18/23"
  sampletable="final_report.csv"
  arpredictions="ar_predictions.tsv"
}
```

```{r prep, echo=FALSE, warning=FALSE}
# read yaml file
config <- read_yaml(configFILE)

## set header text
subHeaderText = paste0(config$sub.title, ": ", reporttype)

## get header table
headerDF <- data.frame(date=date,project=projectname,name=username)

## get summary text
summaryTEXT <- config$summary.paragraph

## get disclaimer text
disclaimerTEXT <- config$disclaimer.text

## get methods text
methodsTEXT <- config$methods.text
```

```{r read, echo=FALSE}
## get sample table
sampleDF <- read.csv2(sampletable,sep=",")
rownames(sampleDF)=sampleDF$specimen_id

# get ar-summary
ar_summary <- read.csv2(arpredictions,sep='\t',check.names = F)
ar_summary$Sample=gsub("_[0-9][0-9][0-9][0-9]_length_[0-9]*","",ar_summary$Sample)
ar_summary$Sample=gsub("_[0-9][0-9][0-9]_length_[0-9]*","",ar_summary$Sample)
ar_summary$Sample=gsub("_[0-9][0-9]_length_[0-9]*","",ar_summary$Sample)
ar_summary$Sample=gsub("_[0-9]_length_[0-9]*","",ar_summary$Sample)
ar_summary$Sample=gsub(paste0("-",projectname),"",ar_summary$Sample)
ar_summary=ar_summary[!duplicated(ar_summary),]

# create metadata
generate_ids=function(list_in){
  tmp_list=which( colnames(sampleDF) %ni% list_in ) -1
  return(tmp_list)
}
id_nums=generate_ids(c("specimen_id","wgs_id","srr_id",
                       "wgs_date_put_on_sequencer","run_id","auto_qc_outcome"))
tax_nums=generate_ids(c("specimen_id","species","sequence_classification",
                        "mlst_1","gamma_beta_lactam_resistance_genes"))
qc_nums=generate_ids(c("specimen_id","estimated_coverage","genome_length",
                        "auto_qc_failure_reason"))
lab_nums=generate_ids(c("specimen_id","species","lab_results"))
sample_meta=SharedData$new(sampleDF)
passed_meta=SharedData$new(subset(sampleDF,auto_qc_outcome=="PASS"))
```

```{r summarytext, echo=FALSE}
# generate summary test for outupt
pass_n=nrow(subset(sampleDF, auto_qc_outcome=="PASS"))
failed_n=nrow(subset(sampleDF, auto_qc_outcome!="PASS"))
failed_samples=subset(sampleDF, auto_qc_outcome!="PASS")$specimen_id
summaryTEXT=paste0("There were ", nrow(sampleDF)," samples analyzed and ",
                   pass_n, " samples passed quality control thresholds, while ",
                   length(failed_samples)," failed.")
if(failed_n>0){
  summaryTEXT=paste0(summaryTEXT,
                     " Failing samples included: ",
                     paste(failed_samples,collapse=" | "),".")
}
```

Information {data-orientation=columns data-icon="fa-info-circle"}
=================================================================
Column {data-width=600}
-----------------------
### `r subHeaderText` 
```{r intro}
# create table
colnames(headerDF) <- c("REPORT DATE","PROJECT NAME","PREPARED BY")

# plot table
pander(headerDF)
```

### METHODS
`r methodsTEXT`
Column {data-width=600}
-----------------------
### Summary
`r summaryTEXT`

### DISCLAIMER
`r disclaimerTEXT`

Datatables {data-orientation=columns data-icon="ion-search"}
=================================================================

Input {.sidebar}
-----------------------------------------------------------------------

### Filters 
```{r filter}
filter_select(
  id="auto_qc_outcome",
  label = "QC Status",
  sharedData = sample_meta,
  group = ~auto_qc_outcome
)
filter_select(
  id="run_id",
  label = "Run ID",
  sharedData = sample_meta,
  group = ~run_id
)
```

Raw {data-width=700 .tabset}
-----------------------------------------------------------------------
### Sample

```{r ids, echo=FALSE}
sample_meta %>%
  DT::datatable(
    extensions = c("Buttons","Scroller"),
    rownames=FALSE,
    style="bootstrap",
    width="100%",
    class="compact",
    options=list(
      dom="Blrtrip",
      deferREander=TRUE,
      columnDefs=list(
        list(
          visible=FALSE,
          targets=id_nums
        )
      ),
      buttons = list(I("colvis"),"csv","excel")
      ),
    colnames = c(
      "ID" = "specimen_id",
      "WGS" = "wgs_id",
      "SRR" = "srr_id",
      "SeqDate" = "wgs_date_put_on_sequencer",
      "RunID" = "run_id",
      "QC" = "auto_qc_outcome"
    )
  )
```

### Taxonomy
```{r tax, echo=FALSE}
sample_meta %>%
  DT::datatable(
    extensions = c("Buttons","Scroller"),
    rownames=FALSE,
    style="bootstrap",
    width="100%",
    class="compact",
    options=list(
      dom="Blrtrip",
      deferREander=TRUE,
      columnDefs=list(
        list(
          visible=FALSE,
          targets=tax_nums
        )
      ),
      buttons = list(I("colvis"),"csv","excel")
      ),
    colnames = c(
      "ID" = "specimen_id",
      "Species" = "species",
      "SeqClass" = "sequence_classification",
      "SeqDate" = "wgs_date_put_on_sequencer",
      "MLST1" = "mlst_1",
      "bLac Genes" = "gamma_beta_lactam_resistance_genes"
    )
  )
```

### QC
```{r qc, echo=FALSE}
sample_meta %>%
  DT::datatable(
    extensions = c("Buttons","Scroller"),
    rownames=FALSE,
    style="bootstrap",
    width="100%",
    class="compact",
    options=list(
      dom="Blrtrip",
      deferREander=TRUE,
      columnDefs=list(
        list(
          visible=FALSE,
          targets=qc_nums
        )
      ),
      buttons = list(I("colvis"),"csv","excel")
      ),
    colnames = c(
      "ID" = "specimen_id",
      "Coverage" = "estimated_coverage",
      "GenomeLength" = "genome_length",
      "QC Fail" = "auto_qc_failure_reason"
    )
  )
```

### ARGene Table
```{r ar_table}
if(exists('ar_summary')){
  
  # determine gene consensus at 50%
  number_of_samples=round(length(unique(ar_summary$Sample))*.5+.5,0)
  output_df=data.frame()
  for (geneID in unique(ar_summary$Gene)){
    tmp_df=subset(ar_summary,Gene==geneID)
    if (nrow(tmp_df)>number_of_samples-1){
      if (nrow(output_df)==0){
        output_df=tmp_df
      } else{
        output_df=rbind(tmp_df,output_df)
      }
    }
  }
  
  # if there are consensus genes then print table
  if (nrow(output_df)==0){
    print("There are no genes with consensus between at least 50% of samples")
  } else{
    # strip rownames, order by gene name
    row.names(output_df) <- NULL
    output_df <- output_df[order(output_df$Gene),]
    DT::datatable(output_df,
          rownames = FALSE,
          extensions = 'RowGroup',
          options = list(rowGroup = list(dataSrc=c(1)),
                         columnDefs = list(list(visible=FALSE, targets=c(1)))
                         )
          )
  }
}
```

### Lab Concordance
```{r lab_conc, echo=FALSE}
sample_meta %>%
  DT::datatable(
    extensions = c("Buttons","Scroller"),
    rownames=FALSE,
    style="bootstrap",
    width="100%",
    class="compact",
    options=list(
      dom="Blrtrip",
      deferREander=TRUE,
      columnDefs=list(
        list(
          visible=FALSE,
          targets=lab_nums
        )
      ),
      buttons = list(I("colvis"),"csv","excel")
      ),
    colnames = c(
      "ID" = "specimen_id",
      "PipelineResults" = "species",
      "LabResults" = "lab_results"
    )
  )
```

Analysis {data-orientation=columns data-icon="ion-android-options"}
=================================================================
Column {data-width=800 .tabset}
-----------------------------------------------------------------------
### Taxonomy
```{r counts}
subset(sampleDF,auto_qc_outcome=="PASS")%>%
ggplot(aes(x = species,fill=species)) +
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### ARGene Summary
```{r ar_summary}
# if there are consensus genes then print table
if (nrow(output_df)==0){
  print("There are no genes with consensus between at least 50% of samples")
} else{
  # update names
  output_df$Sample=gsub(paste0("-",projectname),"",output_df$Sample)

  merge(sampleDF,output_df, by.x="specimen_id",
    by.y="Sample")[,c("specimen_id","Gene","species","run_id")] %>% 
    ggplot(aes(x = Gene,fill=species)) +
    geom_bar() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    facet_grid(~run_id)
}
```

Column {data-width=300}
-----------------------------------------------------------------------
### Notes

#### ARGene
`r config$ar_gene_summary.text`